Robots Showing Emotions: Emotion Representation with no bio-inspired Body

### 2016-10-30 JournalInteractionStudies/Parts/case_studies.tex
Changes in some parts of the article, main changes are described in Rebuttal2.txt

In total, we did one pilot and four case studies using changes in linear and angular velocity, and oscillation angle to convey emotion. These features are depicted in Figure~\ref{fig:features}. The pilot was used to test the prototype and identify possible improvements. The values for the first two case studies were selected empirically, while for the third the Laban's description of emotion expression was followed. The final case study used values identified through the previous trials.

%TODO add some information about the relationship among all the case studies
\begin{figure}[h]
	\centering
	\includegraphics[width=0.75\textwidth]{./Images/ExampleMovement.png}
	\caption{Example of the features used in all case studies. $x$ represents the displacement in meters, $\omega$ is the angular velocity ($rad/s$) and $\theta$ the oscillation of the body around its center ($rad$). The upper sequence depicts a movement based only on linear velocity, while the bottom one shows a sequence with angular and linear movement.}
	\label{fig:features}
\end{figure} 
%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Design}

Each group of participants was exposed to three rounds, in each of which the robot was performing a different emotion expression. The sequence of rounds was generated randomly without repeating any emotion in the same sequence. The subjects were asked to mark which emotion-related term that best described what they believed the robot was trying to convey, selecting among a set of nine possible options: anger, curiosity, disgust, embarrassment, fear, happiness, sadness, neutral, and pride. It was also added the option to answer: ''Unknown''. Although these ten options were enlisted, just the first eight emotions were implemented. To avoid that spectators could be influenced by previous trials, different emotions were shown to each group.

Additionally, it was decided to locate the robot $1.5 mts$ away from the participants and with its front facing them every time that a new emotion was showed, as shown in  Figure~\ref{fig:setup}.

@@

It was decided to follow an empirical design for this case study, consisting in taking into account the features that could be changed in the robot base and in composing them to express some basic emotions. The first implementations where tested and tuned basing on the emotions perceived by the researchers from the robot's movement.\\
The features that could easily be perceived by the observer are summarized in table~\ref{table:features}, and each selected feature is described here below:

@@

\begin{table}[tbh]
\begin{center}
\small
\caption{Features that were presented to the audience, and their values. Embarr. is Embarrassment and Asy. is Asymmetric.}
\label{table:features}
\begin{tabular}{|p{1.5 cm}|p{1.5 cm}|c|p{1.5 cm}|p{1.6 cm}|p{1.6 cm}|m{1.5 cm}|}
\hline 
\textbf{Emotion} & \textbf{Speed} & \textbf{Front/Back} & \textbf{Shoulder} & \textbf{Shoulder Amplitude} & \textbf{Angular Velocity} & \textbf{Body Rotation Amplitude} \\ 
\hline
Anger & Fast & No & Asy. & Medium & None & Very Small \\
\hline
Sadness & Very Slow & Yes-1 & Close-2 & None & Slow & Small \\
\hline
Fear & Normal Slow & Yes-2 & Close-2 + Asy. & Small & None & None \\ 
\hline
%Fear 2 & Slow & Yes-2 & Close-2 + Asymmetric & Small & None & Small \\
%\hline
Embarr. & Slow & No & Asy. & Large & Slow & Small \\
\hline
Happiness & Fast & No & Asy. & Medium & Fast & Small \\
\hline
Disgust & Slow & No & Few Asy. & Small & None & Very Small \\
\hline
Curiosity & Normal & No & Asy. & Medium & Slow & Large \\
\hline
Neutral & Normal & No & Asy. & Medium & None & None \\
\hline
\end{tabular}
\end{center}
\vspace{-0.5cm}
\end{table}

@@

Table~\ref{table:results_1} shows the results obtained in the first case study. 
\begin{table}[tbh]
\caption{Answers obtained the first case study. On each row is the emotion that was intended to express, and on the columns the reported emotions.
.}
\small
\label{table:results_1}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
\backslashbox{Presented}{Reported} & 
\rotatebox{90}{\textbf{Anger}}&
\rotatebox{90}{\textbf{Curiosity}}&
\rotatebox{90}{\textbf{Disgust}}&
\rotatebox{90}{\textbf{Embarr.}}&
\rotatebox{90}{\textbf{Fear}}&
\rotatebox{90}{\textbf{Happiness}}&
\rotatebox{90}{\textbf{Neutral}}&
\rotatebox{90}{\textbf{Pride}}&
\rotatebox{90}{\textbf{Sadness}}&
\rotatebox{90}{\textbf{Unk.}}&
\rotatebox{90}{\textbf{Tot.}}&
\rotatebox{90}{\textbf{Percentage}}\\
\hline
Anger &23 &11 &3 &4 &7 &6 &1 &0 &0 &4 &59&38.98\%\\
\hline
Curiosity &2 &9 &3 &9 &2 &3 &2 &8 &3 &0 &41&21.95\%\\
\hline
Disgust& 1 & 11& 1& 3& 7& 3& 6& 8& 5& 1& 46&2.17\%\\
\hline
Embarr. & 3& 11& 0& 8& 9& 3& 0& 4& 2& 1& 41&19.51\%\\
\hline
Fear & 3& 29& 5& 25& 22& 2& 2& 0& 2& 2& 92&22.82\%\\
\hline
Happiness & 15& 4& 0& 2& 3& 28& 4& 4& 2& 1& 63&44.44\%\\
\hline
Neutral & 5& 0& 1& 1& 0& 1& 0& 3& 2& 0& 13&0\%\\
\hline
Sadness & 4& 5& 1& 11& 13& 0& 2& 3& 7& 0& 46&23.81\%\\
\hline
\end{tabular}
\end{table}

Table~\ref{table:results_1} shows that the best emotion perceived was Happiness and the worst one Disgust, with 44\% and 2\% respectively of accuracy. Moreover, some movements were designed to convey a specific emotion, but a different emotion was perceived. For example the intended Sadness was identified by 14\% of subjects exposed to it, while 25.5\% perceived it as Fear. The neutral behaviour was not much considered, possibly because people, in this setting, expected to see something specific, not a neutral movement. To analyse whether each emotional expression was misinterpreted with others a Fisher's exact test was applied for each possible combination of emotions implemented. Additionally, a Holm-Bonferroni correction was applied for multiple comparisons to get better p-value estimation. 
The results showed that people perceived the following implementations as similar, having a $p-value>0.05$: curiosity-disgust, curiosity-embarrassment, curiosity-sadness, disgust-embarrassment, disgust-sadness, embarrassment-fear, embarrassment-sadness, and fear-sadness.

@@

\begin{table}[h]
\begin{center}
\small
\caption{Example of table compiled for each emotion on the subjects that have been presented each emotion (here Happiness).}
\label{table:singleEmotion}
\begin{tabular}{|c|c|c|}
\hline 
\backslashbox{Presented}{Reported}&Happiness&Other\\
\hline 
Happiness&28&4\\
\hline 
Other&30&96\\
\hline
\end{tabular}
\end{center}
%\vspace{-0.5cm}
\end{table}

@@

The result shows that it is possible to convey some emotions using just movement and a non-bio-inspired embodiment. Moreover, showing emotional states in a context that puts in evidence the stimulus that produced them (as it would happen in a real application) might improve the perception of the emotions, as it is known also for human-human relationships, and this is study in the second case study.

%%%%%%%%%%%%%%%%%%%%%
\section{Second Case Study: Addition of Context}

Since the results of the first case study showed that the recognition of most of the designed emotions was still quite poor, it was decided to try again with the same emotions, but this time a context to the emotional state of the robot was provided by playing a little scene. This is motivated by the fact that that emotions are triggered by stimuli~\cite{Plutchik2001,cacioppo2000handbook}.
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Design}

A little scene was created for each emotion to give a context to the participants. The procedure and questionnaire were the same as the one adopted for the first case. However, it was added a variant of the Fear emotion (Fear2), in which the robot finishes its trajectory far from the person. As done in the first case, each group was exposed to three emotions and the sequence presented to each group was different to eliminate the chance to influence other groups. The sequences were generated randomly and without repeating an emotion in a sequence.

As in the first case study, the total distance covered by the robot was $1.5 mts$, but this time the robot was facing the actor. The distance from the participants to the robot and actor was $1.5 mts$, as shown in Figure~\ref{fig:setup2}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{./Images/SecondCase.png} 
	\caption{Setup of the second case study.}
	\label{fig:setup2}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study}

The laboratory open day of 2013 was used to perform the scenes and actions and to collect data. Eight groups participated, each one with approximately 20 volunteers for a total of 156: 51 males, 17 females, and 88 that did not provide their gender. The average age was 26.34 with a standard deviation of 12.44, and with a minimum age of 11 and maximum of 65.

@@

In the second case study, the results were considerably better, where four out of the nine showed emotions had a percentage of recognition higher than 60\% (Anger, Happiness, Curiosity and Fear2). Another important point to highlight is that Sadness, Fear and Embarrassment were still perceived as different emotions. As it was done for the first case study, a Fisher's exact test and Holm-Bonferroni correction were applied for each possible combination of implemented emotions. The results suggest that no implementation was considered as similar to others by the participants, since for all we have $p-value<\alpha$.  Comparing the percentage from both cases, it is possible to see that giving information about the context that produces the current emotional state of the robot improves its recognition. However, if the conveyed emotion does not match the scene, the audience does not recognize it. This shows that even giving information about the situation, and using a human actor to provide the context, people could not recognize the emotion if the correct features are not exploited. To verify this, it was done a comparison between the results obtained for each emotion in both cases doing a Fisher's exact test. The results suggest that implementations of Disgust ($p=0.088$), Fear ($p=0.206$) and Sadness ($p=0.269$) are perceived as the same in both cases. This would suggest that the combination of stimuli (scene + actor) and correct robot's movements are necessary to the correct emotional interpretation. This opens the question about the influence that could have the human actor and the scene on the recognition rate. Although the study of their influence could be beneficial, it was decided to continue the studies on the values that could be used to express emotions. Therefore it was devise a third case study that uses values informed in the literature.

@@

From the previous experience it was decided to implemented just four emotions, which reduce the number of possible sequences. To select the emotions to be implemented, the \textit{circumplex model of affect} was adopted, as it was suggested by Sharma. The emotions implemented were: \textit{Happiness} (I quadrant), \textit{Anger} (II quadrant), \textit{Sadness} (III quadrant), and \textit{Content} (IV quadrant). To reduce the probability that the correct emotion could be selected by chance four additional emotions were selected following the same selection procedure of the implemented ones, one from each quadrant. These emotions were: \textit{Frustration}, \textit{Boredom}, \textit{Astonishment} and \textit{Tiredness}. The final questionnaire had nine possibilities (eight emotions and ``Unknown'').

In order to study the influence that could be generated by the given list of emotions to the participants, it was used an open questionnaire that did not list any emotion, but rather had a space where participants could write a term to answer the question: "\textit{What emotion the robot has expressed?}".
The setup used in this case study is depicted in the Figure~\ref{fig:setup2}.

@@

Due to the scarce literature about emotional parameters in robotics, the selection of the parameters for the implemented emotions were made following the literature in human emotion projection~\cite{Nadia2013, Crane2013}.
Generally speaking, these works use the Laban's notation to code their findings, which leaves to the designer (and the reader) their specific interpretation.

@@

	\item \textit{\textit{Anger}:} the strong force required by Laban was interpreted as a high speed and then a sudden drop, thus it was decided to have a velocity of $500mm/sec$ for $50mm$ and then a velocity of $300mm/sec$. An oscillation during the whole trajectory was also added; that was between [0.086, -0.086] radians with an angular velocity of $3.5 rad/sec$.

	\item \textit{\textit{Happiness}:} as in \textit{Anger}, for the first $50mm$ the velocity was $500mm/sec$, then it went down to $400mm/sec$ for another $50mm$, and then up again to $800mm/sec$ for the rest of the movement. 
	An oscillation between [0.26, -0.26] radians with an angular velocity of $3rad/sec$ was also added.

	\item \textit{Sadness:} the required "low energy" was interpreted as low speed and then slowing down the pace, thus a velocity of $150mm/sec$ was used for the first $200mm$, a velocity $250mm/sec$ for the next $200mm$, and for the rest a velocity of $300mm/sec$. This emotion expression did not include any angular rotation.

	\item \textit{Content:} the velocity was $150mm/sec$ for the first $200mm$, then $300mm/sec$ for the rest. The maximum and minimum angle of oscillation were [0.175, -0.175] with angular velocity of $0.17rad/sec$
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study}

This case study was done during the Rome Maker Faire, 2014. During a period of four days, people were asked to participate to this study. Each subject was exposed to two rounds: in each one, the robot was performing a different emotion. During the first two days, the open questionnaire was used, while during the last two days the multiple option questionnaire was used. The total number of volunteers for the closed list questionnaire was 91: 52 males, 38 females, and 11 that chose not to specify their gender. The average age was 30.98 years, with standard deviation of 15.12, minimum age was 4 and maximum 71. The total amount of participants assessing the study with the open questionnaire was 84: 47 males, 36 females, and 1 that did not specify; the average age was 24 years, with standard deviation of 15.8, minimum age was 5 and maximum 59.

@@

From the open questionnaire, different terms to describe both emotions and mental states were obtained. In table~\ref{table:open_questionnaire} are reported the results. In order to reduce the quantity of words, words that have a similar meaning were grouped, for example words as \textit{Fear}, \textit{Terror}, \textit{Scared}, and \textit{Worried} were grouped under the label \textit{Fear}. Also, three out of four emotions (\textit{Anger}, \textit{Happiness} and \textit{Sadness}) were at least once listed correctly. \textit{Sadness} and \textit{Content} were mostly perceived as \textit{Fear}, being written 31\% and 28\% of the time for each of these two emotions respectively.

@@

	\item The movements that could be designed as emotional are also associated to mental states of the robots. This is evident in the words obtained from the presentation of four ''emotional'' movements, such as \textit{cold} (referring that the robot was feeling cold), \textit{tenderness}, \textit{shyness}, \textit{power}, \textit{hurry}, \textit{looking for something}, \textit{he wants something}, \textit{enthusiasm}, \textit{sympathy}, \textit{vanity}, and \textit{obedience}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
		\caption{Results obtained from the open questionnaire.}
		\label{table:open_questionnaire}
		\small
			\begin{tabular}{|c|c|c|c|c|c|}
				\hline
					\backslashbox{Presented}{Reported}&\textit{Anger}&\textit{Happiness}&\textit{Sadness}&\textit{Content}&Total\\
				\hline
					\textit{Anger}& 3&6&7&5&21\\
				\hline
					\textit{Happiness}&12& 18&0&6&36\\
				\hline
					\textit{Sadness}&0&0& 8&3&11\\
				\hline
					\textit{Content}&2&0&0& 0&2\\
				\hline
					\textit{Fear}&4&5&14&13&36\\
				\hline
					\textit{Tiredness}&1&0&1&1&3\\
				\hline
					\textit{Cold}&2&0&2&1&5\\
				\hline
					\textit{Shyness}&1&0&4&4&9\\
				\hline
					\textit{Agitated}&4&1&0&0&5\\
				\hline
					\textit{Power}&4&1&0&0&5\\
				\hline
					\textit{Hurry}&0&1&0&0&1\\
				\hline
					\textit{Looking for something}&1&0&1&0&2\\
				\hline
					\textit{He wants something}&1&0&1&0&2\\
				\hline
					\textit{Tenderness}&1&0&1&2&4\\
				\hline
					\textit{Uncertainty}&1&2&0&0&3\\
				\hline
					\textit{Enthusiasm}&1&0&0&0&1\\
				\hline
					\textit{Sympathy}&1&0&2&4&7\\
				\hline
					\textit{Curiosity}&1&2&3&3&9\\
				\hline
					\textit{Shame}&0&0&0&1&1\\
				\hline
					\textit{Nothing}&1&2&1&0&4\\
				\hline
					\textit{Vanity}&1&0&0&0&1\\
				\hline
					\textit{Obedience}&0&0&0&1&1\\
				\hline
					\textit{Total}&43&38&44&46&171\\
				\hline
			\end{tabular}
\end{table}

@@

\begin{table}[h]
\centering
\small
\caption{Answers obtained in the case study.}
		\label{table:result_list_emotions}
		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
			\hline	
\rotatebox{90}{\backslashbox{Presented}{Reported}}&
\rotatebox{90}{\textit{Anger}}&
\rotatebox{90}{ \textit{Happiness}} &
\rotatebox{90}{\textit{Sadness}}&
\rotatebox{90}{\textit{Content}}&
\rotatebox{90}{\textit{Frustration}}&
\rotatebox{90}{\textit{Boredom}}&
\rotatebox{90}{\textit{Astonishment}}&
\rotatebox{90}{\textit{Tiredness}}&
\rotatebox{90}{\textit{Unknown}}&
\rotatebox{90}{Total}\\	
			\hline
				\textit{Anger}& 19&8&1&5&7&2&5&1&0&48\\
			\hline
				\textit{Happiness}&18& 19&0&9&5&1&5&1&0&58\\
			\hline
				\textit{Sadness}&1&9& 12&2&4&5&7&6&0&46\\
			\hline
				\textit{Content}&2&8&7& 4&6&8&5&10&0&50\\	
			\hline	
			\end{tabular}
\end{table}

>>>>

In total, we did one pilot and four case studies using changes in linear and angular velocity, and oscillation angle to convey emotions. These features are depicted in Figure~\ref{fig:features}. The pilot was used to test the prototype and identify possible improvements. The features' values for the first case study were selected empirically and just robots' movement was presented to participants. The results suggested that was possible convey some emotions, but it was decided to give a context to each emotion to verify if that could increased the recognition rate. Therefore, a second case study was done. Each emotion was presented to the participants in a small scene, which was portrayed by the robot and a non professional human actor. The results suggested that showing the emotion in a context improve the recognition of some emotions. However the study of the impact of the actor and context were out of this research. To compare literature on emotion projection, a third case study was done with parameters collected from diverse works, which encoded their emotions' description using Laban's. With similar results between first and third case study, we decided to do an experiment to determine values that better represent some emotions~\cite{Angel2017}. Thus a final case study was done to cross-validate experiment's findings.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.75\textwidth]{./Images/ExampleMovement.png}
	\caption{Example of the features used in all case studies. $x$ represents the displacement in meters, $\omega$ is the angular velocity ($rad/s$) and $\theta$ the oscillation of the body around its center ($rad$). The upper sequence depicts a movement based only on linear velocity, while the bottom one shows a sequence with both angular and linear movement.}
	\label{fig:features}
\end{figure}

@@

\subsection{Experimental Design}

Each group of participants was exposed to three rounds, in each of which the robot was expressing a different emotion. The sequence of rounds was generated randomly without repeating any emotion in the same sequence. The subjects were asked to mark the emotion-related term that best described what they believed the robot was trying to convey, selecting among a set of nine possible options: anger, curiosity, disgust, embarrassment, fear, happiness, sadness, neutral, and pride. The option to answer ``Unknown'' was also included. Although these ten options were enlisted, just the first eight emotions were implemented. To avoid that spectators could be influenced by previous trials, different emotions were shown to each group.

Additionally, every time that a new emotion was showed, the robot started $1.5 mts$ away from the participants, with its front facing them, as shown in  Figure~\ref{fig:setup}.

@@

It was decided to follow an empirical design for this case study, by considering the features that could be changed in the robot base and by composing them to express some basic emotions. The first implementations where tested and tuned basing on the emotions perceived by the researchers from the robot's movement. The features that could be easily perceived by the observer are summarized in table~\ref{table:features}, and each selected feature is described here below:

@@

\begin{table}[tbh]
\begin{center}
\small
\caption{Features that were presented in the first case study to the audience, and their respective values. Embarr. is Embarrassment and Asy. is Asymmetric.}
\label{table:features}
\begin{tabular}{|p{1.5 cm}|p{1.5 cm}|c|p{1.5 cm}|p{1.6 cm}|p{1.6 cm}|m{1.5 cm}|}
\hline 
\textbf{Emotion} & \textbf{Speed} & \textbf{Front/Back} & \textbf{Shoulder} & \textbf{Shoulder Amplitude} & \textbf{Angular Velocity} & \textbf{Body Rotation Amplitude} \\ 
\hline
Anger & Fast & No & Asy. & Medium & None & Very Small \\
\hline
Sadness & Very Slow & Yes-1 & Close-2 & None & Slow & Small \\
\hline
Fear & Normal Slow & Yes-2 & Close-2 + Asy. & Small & None & None \\ 
\hline
%Fear 2 & Slow & Yes-2 & Close-2 + Asymmetric & Small & None & Small \\
%\hline
Embarr. & Slow & No & Asy. & Large & Slow & Small \\
\hline
Happiness & Fast & No & Asy. & Medium & Fast & Small \\
\hline
Disgust & Slow & No & Few Asy. & Small & None & Very Small \\
\hline
Curiosity & Normal & No & Asy. & Medium & Slow & Large \\
\hline
Neutral & Normal & No & Asy. & Medium & None & None \\
\hline
\end{tabular}
\end{center}
\vspace{-0.5cm}
\end{table}

@@

Table~\ref{table:results_1} shows the results obtained in the first case study. 
\begin{table}[tbh]
\caption{Answers obtained in the first case study. On each row is the emotion that was intended to express, and on the columns the reported emotions.}
\small
\label{table:results_1}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
\backslashbox{Presented}{Reported} & 
\rotatebox{90}{\textbf{Anger}}&
\rotatebox{90}{\textbf{Curiosity}}&
\rotatebox{90}{\textbf{Disgust}}&
\rotatebox{90}{\textbf{Embarr.}}&
\rotatebox{90}{\textbf{Fear}}&
\rotatebox{90}{\textbf{Happiness}}&
\rotatebox{90}{\textbf{Neutral}}&
\rotatebox{90}{\textbf{Pride}}&
\rotatebox{90}{\textbf{Sadness}}&
\rotatebox{90}{\textbf{Unk.}}&
\rotatebox{90}{\textbf{Tot.}}&
\rotatebox{90}{\textbf{Percentage}}\\
\hline
Anger &23 &11 &3 &4 &7 &6 &1 &0 &0 &4 &59&38.98\%\\
\hline
Curiosity &2 &9 &3 &9 &2 &3 &2 &8 &3 &0 &41&21.95\%\\
\hline
Disgust& 1 & 11& 1& 3& 7& 3& 6& 8& 5& 1& 46&2.17\%\\
\hline
Embarr. & 3& 11& 0& 8& 9& 3& 0& 4& 2& 1& 41&19.51\%\\
\hline
Fear & 3& 29& 5& 25& 22& 2& 2& 0& 2& 2& 92&22.82\%\\
\hline
Happiness & 15& 4& 0& 2& 3& 28& 4& 4& 2& 1& 63&44.44\%\\
\hline
Neutral & 5& 0& 1& 1& 0& 1& 0& 3& 2& 0& 13&0\%\\
\hline
Sadness & 4& 5& 1& 11& 13& 0& 2& 3& 7& 0& 46&23.81\%\\
\hline
\end{tabular}
\end{table}

Table~\ref{table:results_1} shows that the best emotion perceived was Happiness and the worst one Disgust, with 44\% and 2\% respectively of accuracy. Moreover, some movements were designed to convey a specific emotion, but a different emotion was perceived. For example the intended Sadness was identified by 14\% of subjects exposed to it, while 25.5\% perceived it as Fear. The neutral behaviour was not much considered, possibly because people, in this setting, expected to see something specific, not a neutral movement. To analyse whether each emotional expression was misinterpreted with others a Fisher's exact test was applied for each implemented combination of emotions. Additionally, a Holm-Bonferroni correction was applied for multiple comparisons to get better p-value estimation. 
The results showed that people perceived the following implementations as similar, having a $p-value>0.05$: curiosity-disgust, curiosity-embarrassment, curiosity-sadness, disgust-embarrassment, disgust-sadness, embarrassment-fear, embarrassment-sadness, and fear-sadness.

@@

\begin{table}[h]
\begin{center}
\small
\caption{Example of table compiled for each emotion on the subjects that have been presented each emotion (here Happiness for the first case study).}
\label{table:singleEmotion}
\begin{tabular}{|c|c|c|}
\hline 
\backslashbox{Presented}{Reported}&Happiness&Other\\
\hline 
Happiness&28&4\\
\hline 
Other&30&96\\
\hline
\end{tabular}
\end{center}
%\vspace{-0.5cm}
\end{table}

@@

The result shows that it is possible to convey some emotions using just movement and a non-bio-inspired embodiment. In particular, stronger emotions have been recognized better than milder ones, and emotions typically conveyed by facial expressions (e.g., Disgust) were less recognized. 
%%%%%%%%%%%%%%%%%%%%%
\section{Second Case Study: Addition of Context}

Since the results of the first case study showed that the recognition of most of the designed emotions was still quite poor, it was decided to try again with the same emotion expressions, but this time a context to the emotional state of the robot was provided by playing a little scene. This is motivated by the fact that that emotions are triggered by stimuli~\cite{Plutchik2001,cacioppo2000handbook}.
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Experimental design}

A little scene was created for each emotion were an actor played the stimulus to trigger an emotion\footnote{The scenes could be seen in the following link \url{https://www.youtube.com/watch?v=AXAglJKLwbI}}; for instance, he yelled at the robot that then showed fear. The procedure and questionnaire were the same as those adopted for the first case. However, a variant of the Fear emotion (Fear2) was added, in which the robot finishes its trajectory far from the person. As done in the first case, each group was exposed to three emotions and the sequence presented to each group was different to eliminate the chance to influence other groups. The sequences were generated randomly, without repeating the same emotion in a sequence.

As in the first case study, the total distance covered by the robot was $1.5 m$, but this time the robot was facing the actor. The distance from the participants to the robot and actor was $1.5 m$, as shown in Figure~\ref{fig:setup2}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{./Images/SecondCase.png} 
	\caption{Setup of the second case study.}
	\label{fig:setup2}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental setup}

The laboratory open day of 2013 was used to perform the scenes and to collect data. Eight groups participated, each one with approximately 20 volunteers for a total of 156: 51 males, 17 females, and 88 that did not provide their gender. The average age was 26.34 with a standard deviation of 12.44, and with a minimum age of 11 and maximum of 65.

@@

In the second case study, the results were considerably better, where four out of the nine showed emotions had a percentage of recognition higher than 60\% (Anger, Happiness, Curiosity and Fear2). Another important point to highlight is that Sadness, Fear and Embarrassment were still perceived as different emotions. As it was done for the first case study, a Fisher's exact test and Holm-Bonferroni correction were applied for each possible combination of implemented emotions. The results suggest that no implementation was considered as similar to others by the participants, since for all we have $p-value<(\alpha = 0.05)$. 
Comparing the percentage from both the first two cases, it is possible to see that giving information about the context that produces the current emotional state of the robot improves the recognition of some emotions, but for others remains still low. Our hypothesis is that there must be a match between robot's movements with the given context and actors performance to increase the recognition rate. If one or all do not fit participants' idea of the emotion, this one is not going to be recognized. To verify this, a comparison between the results obtained for each emotion in both cases doing a Fisher's exact test was done. The results suggest that implementations of Disgust ($p=0.088$), Fear ($p=0.206$) and Sadness ($p=0.269$) are perceived as the same in both cases, which correspond to the emotions with lower recognition rate. This opens the question about the influence that could have the human actor and the scene on the recognition rate. Although the study of this could be beneficial, it was decided to leave this to future work. Instead, a third case study was devised, where values presented in the literature for emotion expression in humans were adopted in the quest for more recognizable emotion expressions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental design}

In order to reduce the number of possible sequences, it was decided to implemented only four emotions. To select the emotions to be implemented, the \textit{circumplex model of affect} was adopted, as it was suggested by Sharma. The emotions implemented were: Happiness (I quadrant), Anger (II quadrant), Sadness (III quadrant), and Content (IV quadrant). To reduce the probability that the correct emotion could be selected by chance four additional emotions were selected following the same selection procedure of the implemented ones, one from each quadrant. These emotions were: Frustration, Boredom, Astonishment and Tiredness. The final questionnaire had nine possibilities (eight emotions and ``Unknown''), and was better founded on psychological theory than the previous ones.

In order to study the influence that could be generated by the given list of emotions to the participants, it was used an open questionnaire that did not list any emotion, but rather had a space where participants could write a term to answer the question: "\textit{What emotion has the robot expressed?}".
The setup used in this case study is depicted in Figure~\ref{fig:setup2}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Emotion Description}

Due to the scarce literature about emotional parameters in robotics, the selection of the parameters for the implemented emotions were made following the literature in human emotion projection~\cite{Nadia2013, Crane2013}.
Generally speaking, these works use the Laban's model to code their emotional expressions, which is based on linguistic terms and leaves to the designer (and the reader) their specific interpretation.

@@

	\item Anger: the strong force required by Laban was interpreted as a high speed and then a sudden drop, thus it was decided to have a velocity of $500mm/sec$ for $50mm$ and then a velocity of $300mm/sec$. An oscillation during the whole trajectory was also added; that was between [0.086, -0.086] radians with an angular velocity of $3.5 rad/sec$.

	\item Happiness: as in Anger, for the first $50mm$ the velocity was $500mm/sec$, then it went down to $400mm/sec$ for another $50mm$, and then up again to $800mm/sec$ for the rest of the movement. 
	An oscillation between [0.26, -0.26] radians with an angular velocity of $3rad/sec$ was also added.

	\item Sadness: the required "low energy" was interpreted as low speed and then slowing down the pace, thus a velocity of $150mm/sec$ was used for the first $200mm$, a velocity $250mm/sec$ for the next $200mm$, and for the rest a velocity of $300mm/sec$. This emotion expression did not include any angular rotation.

	\item Content: the velocity was $150mm/sec$ for the first $200mm$, then $300mm/sec$ for the rest. The maximum and minimum angle of oscillation were [0.175, -0.175] with angular velocity of $0.17rad/sec$
\end{itemize}

@@

This case study was done during the Rome Maker Faire, 2014. During a period of four days, people were asked to participate to this study. Each subject was exposed to two rounds: in each one, the robot was performing a different emotion. During the first two days, the open questionnaire was used, while during the last two days the multiple option questionnaire was used. The total number of volunteers for the closed list questionnaire was 91: 52 males, 38 females, and 11 that did not specify their gender. The average age was 30.98 years, with standard deviation of 15.12, minimum age was 4 and maximum 71. The total amount of participants assessing the study with the open questionnaire was 84: 47 males, 36 females, and 1 that did not specify; the average age was 24 years, with standard deviation of 15.8, minimum age was 5 and maximum 59.

@@

From the open questionnaire, different terms to describe both emotions and mental states were obtained. In table~\ref{table:open_questionnaire} are reported the results. In order to reduce the quantity of terms, words that have a similar meaning were grouped, for example words such as Fear, Terror, Scared, and Worried were grouped under the label Fear. Also, three out of four emotions (Anger, Happiness and Sadness) were at least once listed correctly. Sadness and Content were mostly perceived as Fear, being reported 31\% and 28\% of the time, respectively.

@@

	\item The movements that could be designed as emotional are also associated to mental states of the robots. This is evident in the words obtained from the presentation of four ``emotional'' movements, such as: cold (referring that the robot was feeling cold), tenderness, shyness, power, hurry, looking for something, he wants something, enthusiasm, sympathy, vanity, and obedience.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
		\caption{Results obtained from the open questionnaire in the third case study.}
		\label{table:open_questionnaire}
		\small
			\begin{tabular}{|c|c|c|c|c|c|}
				\hline
					\backslashbox{Presented}{Reported}&Anger&Happiness&Sadness&Content&Total\\
				\hline
					Anger& 3&6&7&5&21\\
				\hline
					Happiness&12& 18&0&6&36\\
				\hline
					Sadness&0&0& 8&3&11\\
				\hline
					Content&2&0&0& 0&2\\
				\hline
					Fear&4&5&14&13&36\\
				\hline
					Tiredness&1&0&1&1&3\\
				\hline
					Cold&2&0&2&1&5\\
				\hline
					Shyness&1&0&4&4&9\\
				\hline
					Agitated&4&1&0&0&5\\
				\hline
					Power&4&1&0&0&5\\
				\hline
					Hurry&0&1&0&0&1\\
				\hline
					Looking for something&1&0&1&0&2\\
				\hline
					He wants something&1&0&1&0&2\\
				\hline
					Tenderness&1&0&1&2&4\\
				\hline
					Uncertainty&1&2&0&0&3\\
				\hline
					Enthusiasm&1&0&0&0&1\\
				\hline
					Sympathy&1&0&2&4&7\\
				\hline
					Curiosity&1&2&3&3&9\\
				\hline
					Shame&0&0&0&1&1\\
				\hline
					Nothing&1&2&1&0&4\\
				\hline
					Vanity&1&0&0&0&1\\
				\hline
					Obedience&0&0&0&1&1\\
				\hline
					Total&43&38&44&46&171\\
				\hline
			\end{tabular}
\end{table}

@@

\begin{table}[h]
\centering
\small
\caption{Answers obtained from the closed list group in the third case study.}
		\label{table:result_list_emotions}
		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
			\hline	
\rotatebox{90}{\backslashbox{Presented}{Reported}}&
\rotatebox{90}{Anger}&
\rotatebox{90}{ Happiness} &
\rotatebox{90}{Sadness}&
\rotatebox{90}{Content}&
\rotatebox{90}{Frustration}&
\rotatebox{90}{Boredom}&
\rotatebox{90}{Astonishment}&
\rotatebox{90}{Tiredness}&
\rotatebox{90}{Unknown}&
\rotatebox{90}{Total}\\	
			\hline
				Anger&19&8&1&5&7&2&5&1&0&48\\
			\hline
				Happiness&18&19&0&9&5&1&5&1&0&58\\
			\hline
				Sadness&1&9&12&2&4&5&7&6&0&46\\
			\hline
				Content&2&8&7&4&6&8&5&10&0&50\\	
			\hline	
			\end{tabular}
\end{table}
